Routines and Regret: An Examination of Behavioral Norms and Emotional Responses

### 2024-02-15 paper/paper.qmd
modify

[@kahnemanMiller1986] emphasized normality as a central determinant of regret and counterfactual thinking-imagining alternative versions of the past about what might have happened differently [@byrne2016; @millerTurnbullMcFarland1990; @roese1997]. They summarize evidence that anomalous behavior tends to trigger more counterfactual thinking and stronger feelings of regret. They discussed the concept of normality extensively, covering experimental studies involving the odds of influencing the perception and/or interpretation of what is normal as a reference point for comparing behaviors to assess whether a person's behavior is normal or exceptional for that person.Lucas Kutschera and Gilad Feldman simulated abnormal behavior and regret through three scenarios and found the correlation between routines and regret. We replicated their paper with expanded applications and practical implications of the research.

@@

## Regret in Unconventional Choices 
For part1, we reproduce and extend a classic experiment on the hitchhiking situation to prove stronger regret for exception than for routines. (reference)We used a survey to collect the participants' answers to different dimensions in this situation, according to regret, norm, and negative impact, and compared the number of answers, and made a table and a bar chart to reflect the proportion of participants' different views. We found that in questions 1, 2, and 4, more than 90% of participants who chose to represent an unconventional situation experienced a higher sense of regret, and in question 3, more than 90% of participants felt that Smith, who represented a regular situation, was more socially normative. The results of questions 2 and 3 support our hypothesis that frequent free-riding behavior is perceived by the general public as less common and is vulnerable to criticism if it conflicts with society's prohibitive norms, while the results of questions 1 and 4 are consistent with our expectation that most participants will have a stronger sense of regret for Jones and more negative feelings and effects later on. This proves the conclusion that regret for exceptional cases will be stronger than regret for normal cases, and the data supports anomalous effects.

@@

We replicated two parts of the (2021) paper-the hitchhiking scenario and the car accident scenario. Overall, through participant data analyses of regret, luck, social norms, and negative affect in different scenarios, we all gained support for estimand: people feel more regret when the event occurs in exception scenarios than when they are harmed in routine events.

>>>>

[@kahnemanMiller1986] emphasized normality as a central determinant of regret and counterfactual thinking-imagining alternative versions of the past about what might have happened differently [@byrne2016; @millerTurnbullMcFarland1990; @roese1997]. They summarize evidence that anomalous behavior tends to trigger more counterfactual thinking and stronger feelings of regret. They discussed the concept of normality extensively, covering experimental studies involving the odds of influencing the perception and/or interpretation of what is normal as a reference point for comparing behaviors to assess whether a person's behavior is normal or exceptional for that person.Lucas Kutschera and Gilad Feldman simulated abnormal behavior and regret through three scenarios and found the correlation between them. We replicated their paper with expanded applications and practical implications of the research.

@@

## Regret in Unconventional Choices 
For scenario 1, we reproduce and extend a classic experiment on the hitchhiking situation to prove stronger regret for exception than for routines. (reference)We used a survey to collect the participants' answers to different dimensions in this situation, according to regret, norm, and negative impact, and compared the number of answers, and made a table and a bar chart to reflect the proportion of participants' different views. We found that in questions 1, 2, and 4, more than 90% of participants who chose to represent an unconventional situation experienced a higher sense of regret, and in question 3, more than 90% of participants felt that Smith, who represented a regular situation, was more socially normative. The results of questions 2 and 3 support our hypothesis that frequent free-riding behavior is perceived by the general public as less common and is vulnerable to criticism if it conflicts with society's prohibitive norms, while the results of questions 1 and 4 are consistent with our expectation that most participants will have a stronger sense of regret for Jones and more negative feelings and effects later on. This proves the conclusion that regret for exceptional cases will be stronger than regret for normal cases, and the data supports anomalous effects.

@@

We successfully and replicated and expanded both parts of the (2021) paper - the hitchhiking scenario and the car accident scenario. By experimentally obtaining data and summarizing and analyzing the data obtained, we concluded that individuals tend to develop higher regret in particular situations in contrast to past normative behaviors. This conclusion provides insight into the psychological problems of people in society and informs new policies to improve the mental health of the population. Also on a personal level it can help individuals to manage their decisions and resolve psychological regrets. In conclusion, by analyzing participant data on regret, luck, social norms, and negative emotions in different scenarios, we all gained support for the estimate that people feel more regret when an event occurs in an exception scenario than when they are harmed in a routine event.
